---
title: "Stepwise Evaluation of Real Estate Data"
output:
  html_document: default
  html_notebook: default
---

##Data Cleaning and Stepwise Evaluation

###Data Cleaning`

We have our intial data cleaning function.

```{r}
  setwd("~/")
  data = read.csv("./train.csv")
  data = data[-1]
  
  data$MSSubClass = as.factor(data$MSSubClass)#Factor encoded as a numeric value
  
  #Median or 0
  data$LotFrontage[is.na(data$LotFrontage) == TRUE] = median(data$LotFrontage[!is.na(data$LotFrontage)])
  #data$LotFrontage[is.na(data$LotFrontage) == TRUE] = median(data$LotFrontage[!is.na(data$LotFrontage)])
  
  data$Alley = factor(data$Alley, levels=c(NA, "None", "Grvl", "Pave"))
  data$Alley[is.na(data$Alley)] = "None"
  
  data$MasVnrType[is.na(data$MasVnrType) == TRUE] = "None"
  
  data$MasVnrArea[is.na(data$MasVnrArea) == TRUE] = 0
  
  data$BsmtQual = factor(data$BsmtQual, levels=c("None", levels(data$BsmtQual)))
  data$BsmtQual[is.na(data$BsmtQual)] = "None"
  
  data$BsmtCond = factor(data$BsmtCond, levels=c("None", levels(data$BsmtCond)))
  data$BsmtCond[is.na(data$BsmtCond)] = "None"
  
  data$BsmtExposure = factor(data$BsmtExposure, levels=c("None", levels(data$BsmtExposure)))
  data$BsmtExposure[is.na(data$BsmtExposure)] = "None"
  
  data$BsmtFinType1 = factor(data$BsmtFinType1, levels=c("None", levels(data$BsmtFinType1)))
  data$BsmtFinType1[is.na(data$BsmtFinType1)] = "None"
  
  data$BsmtFinType2 = factor(data$BsmtFinType2, levels=c("None", levels(data$BsmtFinType2)))
  data$BsmtFinType2[is.na(data$BsmtFinType2)] = "None"
  
  data$Electrical[is.na(data$Electrical) == TRUE] = "SBrkr"
  
  data$FireplaceQu = factor(data$FireplaceQu, levels=c("None", levels(data$FireplaceQu)))
  data$FireplaceQu[is.na(data$FireplaceQu)] = "None"
  
  data$GarageType = factor(data$GarageType, levels=c("None", levels(data$GarageType)))
  data$GarageType[is.na(data$GarageType)] = "None"
  
  data$GarageYrBlt[is.na(data$GarageYrBlt) == TRUE] = 0
  
  data$GarageFinish = factor(data$GarageFinish, levels=c("None", levels(data$GarageFinish)))
  data$GarageFinish[is.na(data$GarageFinish)] = "None"
  
  data$GarageQual = factor(data$GarageQual, levels=c("None", levels(data$GarageQual)))
  data$GarageQual[is.na(data$GarageQual)] = "None"
  
  data$GarageCond = factor(data$GarageCond, levels=c("None", levels(data$GarageCond)))
  data$GarageCond[is.na(data$GarageCond)] = "None"
  
  data$PoolQC = factor(data$PoolQC, levels=c("None", levels(data$PoolQC)))
  data$PoolQC[is.na(data$PoolQC)] = "None"
  
  data$Fence = factor(data$Fence, levels=c("None", levels(data$Fence)))
  data$Fence[is.na(data$Fence)] = "None"
  
  data$MiscFeature = factor(data$MiscFeature, levels=c("None", levels(data$MiscFeature)))
  data$MiscFeature[is.na(data$MiscFeature)] = "None"
  


```

Some terms are added and cleaned for clarity's sake. Quality variables are broken down to a baseline of 5(average) where a positive term will indicate improvement above average, and a negative will be below average

Year variables are replaced as their distance from 2000, rather than from year 0.


```{r}
#Changing Base 10 variables so they are representative of change from the average
data$OverallQual= (data$OverallQual - 5)
data$OverallCond= (data$OverallCond - 5)

#Changing the base of  year variables so they are based off of distance from 2000
data$YearBuilt= data$YearBuilt - 2000
data$YearRemodAdd= data$YearRemodAdd - 2000
data$GarageYrBlt= data$GarageYrBlt - 2000
data$YrSold = data$YrSold -2000

```


We create some binary variables which will add more precision in explaining the data.
```{r}
data$CentralAir<-data$CentralAir == "Y"

data$HasBsmt <- data$BsmtCond != "None"

data$Has2ndFlr<-data$X2ndFlrSF != 0 

#We want Fireplaces as a factor instead of an integer
data$Fireplaces <- factor(data$Fireplaces)

data$HasGarage <- data$GarageCond != "None"

data$HasPool <- data$PoolQC != "None"

data$HasFence <- data$Fence != "None"

data$GarageCars<- factor(data$GarageCars)

data$KitchenAbvGr <- factor(data$KitchenAbvGr)

data$HasWoodDeck <- data$WoodDeckSF != 0
data$HasOpenPorch<- data$OpenPorchSF != 0
data$HasEnclosedPorch<- data$EnclosedPorch != 0
data$Has3SsnPorch <- data$X3SsnPorch != 0

```


We change Sale Price to a log variable so that our coefficients can be interpreted as percentages. 
```{r}

data$SalePrice<- log(data$SalePrice)
```


###Stepwise Data Evaluation
We increase the number of steps, incase it provides a more precise value
```{r}
x<-data[-80]
y<-data[80]
#Stepwise Regression
full<-lm(y$SalePrice~ ., data = x )
coef.back<-step(full, direction = "backward", trace = FALSE, steps=5000)
coef.for<-step(full, direction = "forward", trace = FALSE, steps=5000)
coef.both <- step(full, direction="both", trace = FALSE, steps=5000)

```

```{r}
#turnoff scientific notation 
options(scipen=999)
summary(coef.back)
length(coefficients(coef.back))

```
###The following variables have been determined significant in backwards elimination 

These can be difficult to see given all the factor variables. 

1.MSSubClass
2. MSZoning
3. Lot Frontage
4. Lot Area
5. Street Pave
6. Land Contour 
7. Utilities
8. Lot Config
9. Land Slope
10. Neighborhood
11. Condition
12. Overall Quality
13. Overall Condition
14. Year Built
15. Year Remodeled
16. Roof Style 
17. Roof Material
18. Exterior Type
19. MasVnrArea
20. External Condition
21. Foundation
22. Basement Quality
23. Basement Condition
24. Basement Exposure
25. Basement Finish
26. Heating Type
27. Central Air
28. 1st Flr Sq Ft
29. 2nd Flr Sq Ft
30. Low QUality Finish
31. Basement FUll Bath
32. Full Bath Number
33. Half Bath Number
34. Kitchen Above Ground
35. Kitchen Quality
36. TOtal Rooms Above Ground 
37. Functional
38. FirePlaces
39. Garage Yr Blt
40. Garage Area
41. Garage Condition
42. Garage Quality
43. WoodDeck SF
44. Enclosed Porch
45 3 Season Porch
46. Screen Porch
47. Pool Area
48. Pool Quality
49. Fence Type
50. Sale Type
51. Sale Condition
52. HasOpenPorch 

So backwards elimination eliminated roughly 27 variables 

Has a R Square of .9425, Adjusted R Square of .9337
Has 193 Terms (Including Factors)

```{r}
summary(coef.for)
```
### Unique Variables in the Set From Backwards Elimination
1. Alley Type
2. Lot Shape
3. Building Type
4. House Style
5. 2nd Exterior Type
6. Exterior Quality
7. Basement FInished SF
8. Basement FInished SF 2
9. Basement Unfinished SF
10. Electrical
11. Basement Half Bath
12. Bedroom Above Ground
13. Fireplace Quality
14. Garage Type
15. Garage Finish
16. Garage Cars
17. Paved Drive
18. Open Porch Sfs
19. MIsc Feature
20. MoSold
21. Yr Sold
22. Has2ndFlr
23.HasWoodDeck
24.HasEnclosedPorch
25.Has3SsnPorch

Has an R square of .9459, Adjusted R square .9334
273 Terms

```{r}
#summary(coef.both)
```
This evaluation returned the same results as backwards selection. 

So we see that forward elimination evaluation explains roughly .2% more of the variation in the data, than backwards selection, but it has roughly 80 more terms (including factors), which equates to the evaluation being roughly 40% bigger for .2% increase in accuracy. 

The largest effects in the backwards elimination model are 
1. MS Subclass 160
2. MS Zoning (all terms)
3. Lot Area
4. Land Slope Severe 
5. StoneBridge Neighborhood
6. Normal Condition
7. Condition2PosN
8. Overall Quality
9. Overall Conidtion
10. Year Built 
11. Year Remodeled
12. Roof material (Huge Effect)
13. Basement Finished SF 1 & 2
14. Basment Unifinished SF
15. Heating QCTA
16. Central Air
17. 1st Floor SF 
18. 2nd Floor SF
19. Significant Effect for Kitchen Quality Below Excellent
20. Functional MAJ 2
21. 1 Fireplace
22. Garage Area
23. Wood Deck SF
24. Screen Porch
25. Sale Condition Normal 


It appears that all the values we selected for in the Random Forests Analysis show up in the stepwise regression, however they show up in a much smaller scale. 





